### 2004-06-29 Documentation/part2.tex
*** empty log message ***

\part{The Abstract Experiment Definition Language -- AEDL}

@@

In order to implement a experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', which is the link between 
the oscillation physics and the detection properties for a specfific oscillation pattern. A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

The channels are the building blocks for the so-called ``rules''. In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies. The event numbers from these channels are added {\em before} the $\chi^2$-value is calculated. In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\chi^2$-value, and the total $\chi^2$ is obtained by adding the $\chi^2$'s of all rules. 
 An example for a rule could look like this: We want to detect electron neutrino appearance (``signal''), where the overall efficiency for quasi-elastics electron neutrino events is $0.4$. There is a fraction of $0.01$ of all neutral current events which are mis-identified as quasi-elastic electron neutrino events (``background''). The neutral current fraction is only known within $10\%$ (``background uncertainty'') and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\chi^2$ which properly includes 
systematical errors. The resulting $\chi^2$ is then the starting point for the
oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule the event numbers are added.
\item
 Within each rule the systematics is treated independently from the other rules.
\item
 For each rule the $\chi^2$ is computed; the $\chi^2$'s from all rules are added.
\end{itemize}

>>>>

\part{The Abstract Experiment Definition Language -- AEDL}
\label{part:2}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=13cm]{AEDL1}
\end{center}
\caption{\label{fig:channel} General concept of a ``channel''.}
\end{figure}

In order to implement a experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', which is the link between 
the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}). A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

\begin{figure}[t]
\begin{center}
\includegraphics[width=8cm]{SignalBackground}
\end{center}
\caption{\label{fig:rule} General concept of a ``rule''.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=9cm]{Rules}
\end{center}
\caption{\label{fig:experiment} General concept of an ``experiment''.}
\end{figure}

The channels are the building blocks for the so-called ``rules''. In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
in this manual, we $\chi^2$ and $\Delta \chi^2$ are equal, since
for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}). 
 An example for a rule could look like this: We want to detect electron
  neutrino appearance (``signal''), where the overall efficiency for 
  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
  $0.01$ of all neutral current events which are mis-identified as 
  quasi-elastic electron neutrino events (``background''). The neutral 
  current fraction is only known within $10\%$ (``background uncertainty'') 
  and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors. The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule the event numbers are added.
\item
 Within each rule the systematics is treated independently from the other rules.
\item
 For each rule the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
\end{itemize}
